
---
title: "Recreating White's Robust Standard Errors"
author: "Marie Hogan"
date: "7/18/2023"
output: github_document
---

```{r setup, include=FALSE, comment=""}
knitr::opts_chunk$set(echo = TRUE)
```

I recreate the robust standard errors provided by R package __car__, using the _Chirot_ dataset preloaded on the companion __carData__ package. I use White's formula for robust standard errors.

```{r cars, echo=TRUE, message=FALSE, warning=FALSE}
#installing necessary packages
#requires package car, lmtest
library(car)
library(lmtest)
library(carData)

#sets up tester dataset, part of carData/car package preloaded data
df <- Chirot
```

**Regression Model**  


This model is for demonstration purposes only and I do not offer any interpretation. Its unit of observation is each county in Romania. There are 32 observations. Outcome variable intensity describes conflict intensity during Peasant Rebellion. For more information on the data, search "Chirot" in the R Help view after installing the package carData.
$$ Conflict Intensity = \alpha + \beta_1 Commerce + \beta_2 Tradition + \beta_3Peasantry + \beta_4Inequality$$

```{r , comment=""}
#manual coefficients
#define x, y
x <- as.matrix(cbind(1,df[2:5]))
y <- as.matrix(df[1])

#estimate coefficients
beta_hat <- solve(crossprod(x))%*%t(x)%*%y
labels = labels(beta_hat)[[1]]
as.data.frame(beta_hat, row.names = c("intercept", labels[2:5]))
```

**Robust Standard Errors**  



Next, I follow White's formula for heteroskedasticity robust standard errors to find standard errors for the estimated coefficients. The formula I follow is below:

$$ (X'X)^{-1}Xdiag(e_i^2)X'(X'X)^{-1}$$

```{r , comment=""}
manual_se = sqrt(diag(solve(crossprod(x))%*%(t(x)%*%(diag(as.vector((y - x%*%beta_hat)^2)))%*%x)%*%(solve(crossprod(x)))))

results = data.frame(beta_hat,manual_se, row.names = c("intercept", labels[2:5]))
colnames(results) = c("manual_coef", "manual_SE")
results

```

**Comparing Against R's Packages**  



Using the strategy recommended in [Principles of Econometrics with R](https://bookdown.org/ccolonescu/RPoE4/heteroskedasticity.html#heteroskedasticity-consistent-standard-errors) I calculate coefficient estimates and robust standard errors using pre-built functions in the __car__ and __lmtest__ packages.
```{r , comment=""}
#built in to check
model = lm(intensity ~ commerce + tradition + midpeasant + inequality, df)
vcov_mat = hccm(model, type = "hc0")
built_in = coeftest(model, vcov = vcov_mat)

```
__Comparing Results__
```{r , comment=""}
#packages the outputs together to view and compare
results$built_in_SE = built_in[,2]
results$built_in_coef = built_in[,1]
results[,c(1,4,2,3)]
```

